---
title: "Numerical features imputation"
author: "Mohamed Moukhtar / KHALIL"
date: "2023-05-19"
output: html_document
---
---
title: "DM3-IMPUTE EPCV"
author: "Mohamed Moukhtar / KHALIL"
date: "2023-04-22"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Exploring the Data
```{r}
library(tidyverse)
library(haven)
library(VIM)
library(mice)
```
```{r}
df = read_sav("C:/Users/USER/Desktop/Base_EPCV2019-2020/Base_EPCV2019-2020/menage_2019.sav")
head(df)
```

The dataframe contains variables with missing values.
```{r}
na_df = df %>%
  summarise_all(~sum(is.na(.))) %>% # count NAs for each column
  gather() %>% # convert to long format
  arrange(value)
#drop(na_df[,c('G11B1','G11B2')])
na_df = as.data.frame(na_df)
na_df

```

```{r}
ggplot(na_df[3:10,], aes(x = key, y = value)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  xlab("Variable") +
  ylab("Number of missing values") +
  ggtitle("Number of missing values per variable")
```

We can see that the number of missing values for each variable is significant.

Begin Imputation
We have selected 5 quantitative variables: F2, F6, I12, I13, and SA6_C.

## Impute F2 - How many rooms are there in your accommodation?
Create a sub-dataframe without missing values of F2 and create new arbitrary missing values.

```{r}
df_notna <- df[!is.na(df$F2), ]
set.seed(123)
sample_indices <- sample(1:nrow(df_notna), round(0.4*nrow(df_notna)), replace=FALSE)
F2_na = df_notna['F2']
F2_na = data.frame(F2_na)
F2_na[sample_indices,] = NA
df_notna = data.frame(df_notna)
df_notna['F2'] = F2_na
```
Imputation evaluation using mice:

```{r}

df_mice <- mice(df_notna[,c(34,35,36,37)], m=3,method = "pmm")

```
Evaluate:
```{r}
plot(density(df[!is.na(df$F2), ]$F2,bw=2), main = "Density Comparison of mice", col = "blue", lwd = 2, ylim = c(0, 1))

# Add a density plot for F6_na
lines(density(as.numeric(unlist(df_mice$imp$F2)),bw=2), col = "red", lwd = 2)
```

Imputation d'evaluation par : k-nn

```{r}
df_knn=kNN(df_notna,variable = 'F2',k=3)

```
Evaluate

```{r}
plot(density(df[!is.na(df$F2), ]$F2), main = "Density Comparison of K-nn", col = "blue", lwd = 2, ylim = c(0, 1))

# Add a density plot for F6_na
lines(density(df_knn$F2), col = "red", lwd = 2)

```

Imputation evaluation using hotdeck:

```{r}
df_hd = hotdeck(df_notna)

```
Evaluate:

```{r}
plot(density(df[!is.na(df$F2), ]$F2), main = "Density Comparison of hotdeck", col = "blue", lwd = 2, ylim = c(0, 1))

# Add a density plot for F6_na
lines(density(df_hd$F2), col = "red", lwd = 2)

```

## Imputing F6 - How many hectares of agricultural land does the household use that it does not own?
Create a sub-dataframe without missing values of F6 and create new arbitrary missing values.

```{r}
df_notna <- df[!is.na(df$F6), ]
set.seed(123)
sample_indices <- sample(1:nrow(df_notna), round(0.4*nrow(df_notna)), replace=FALSE)
F6_na = df_notna['F6']
F6_na = data.frame(F6_na)
F6_na[sample_indices,] = NA
df_notna = data.frame(df_notna)
df_notna['F6'] = F6_na
```

Imputation evaluation using mice:
```{r}

df_mice <- mice(df_notna[,c(39,40)], m=3,method = "pmm")

```
Evaluation:
```{r}
plot(density(df[!is.na(df$F6), ]$F6), main = "Density Comparison of mice", col = "blue", lwd = 2, ylim = c(0, 1))

# Add a density plot for F6_na
lines(density(as.numeric(unlist(df_mice$imp$F6))), col = "red", lwd = 2)
```

Imputation evaluation using k-nn:

```{r}
df_knn=kNN(df_notna,variable = 'F6',k=3)

```
Evaluation:

```{r}
plot(density(df[!is.na(df$F6), ]$F6), main = "Density Comparison of knn", col = "blue", lwd = 2, ylim = c(0, 1))

# Add a density plot for F6_na
lines(density(df_knn$F6), col = "red", lwd = 2)

```

Imputation evaluation using hotdeck:

```{r}
df_hd = hotdeck(df_notna)

```
Evaluation:

```{r}
plot(density(df[!is.na(df$F6), ]$F6), main = "Density Comparison of hotdeck", col = "blue", lwd = 2, ylim = c(0, 1))

# Add a density plot for F6_na
lines(density(df_hd$F6), col = "red", lwd = 2)

```


